---
title: "2022 US Political Survey Analysis"
author: 
  - Yimiao Yuan
thanks: "Code and data are available at: https://github.com/YimiaoYuan09/US_Election_2022"
date: today
date-format: long
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(rstanarm)
library(modelsummary)
library(knitr)
library(marginaleffects)
library(tibble)
library(margins)
```


# Introduction
In this paper, I use the 2022 Cooperative Election Study (CES) [@citeCes2022] as our primary dataset. This is a long-established annual political survey of US. In 2022, there were 61,000 American adults completed the survey. Since it is difficult to observe such a large dataset, this paper will only analyze respondents who are registered to vote and focus on Biden and Trump's results. 

The dataset is downloaded, cleaned and analyzed using statistical programming language R [@citeR], with packages `tidyverse` [@citeTidyverse], `here` [@citeHere], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModleSummary], `knitr` [@citeKnitr], `marginaleffects` [@citeMarginalEffects].


```{r}
#| include: false
#| warning: false
#| message: false

# read in cleaned data
ces2022 <-  read_csv(here::here("data/analysis_data/cleaned_ces2022.csv"))
```

```{r}
#| include: false
#| warning: false
#| message: false

# change column type to factor
ces2022 <-
  ces2022 |>
  mutate(
    voted_for = as_factor(voted_for),
    race = factor(
      race,
      levels = c(
        "White",
        "Black",
        "Hispanic",
        "Asian",
        "Native American",
        "Middle Eastern",
        "Two or more races",
        "Other"
      )
    )
  ) |>
  select(voted_for, race)
```


# Model {#sec-model}
The models are created to analyze the relationship between voter preferences and voter's race. In the model, $y_i$ represents whether an individual voted Biden or Trump. For all parameters, I use a normal distribution with a mean of 0 and a standard deviation of 2.5.

## Logistic Regression

\begin{align} 
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \alpha + \beta_1 \times \mbox{race}_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5)
\end{align}

## Poisson Regression

\begin{align} 
y_i|\lambda_i &\sim \mbox{Poisson}(\lambda_i) \\
\mbox{logit}(\lambda_i) &= \alpha + \beta_1 \times \mbox{race}_i\\
\end{align}

## Model Results

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: tbl-modelresults1
#| tbl-cap: "Logistic and Poisson model results"

logistic_model <-
  readRDS(file = here::here("models/logistic_model.rds"))

poisson_model <-
  readRDS(file = here::here("models/poisson_model.rds"))


modelsummary(
  list(
    "Logistic" = logistic_model,
    "Poisson" = poisson_model
  )
)
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: tbl-modelresults2
#| tbl-cap: "Negative Binomial model results"

neg_bio_model <-
  readRDS(file = here::here("models/neg_bio_model.rds"))

summary(neg_bio_model)[1:11, 1:3] |> 
  kable()
```



The results are summarized in @tbl-modelresults1 and @tbl-modelresults2. Poisson and Negative Binomial model have similar results, while Logistic model has a slightly different one. Logistic model has an intercept of around 1.7, but poisson and negative binomial model is about -0.2. Other parameters have similar trends in these three models, but with different values.


## Posterior predictive check
@fig-post_dist shows the posterior distribution for three models. According to the graphs, I can conclude that Logistic Regression is a better choice for current situation. Also, since our outcome (vote for Biden/Trump) is a binary variable, Logistic regression fits better. Poisson regression and negative binomial regression are used where the dependent variable represents the count of events occurring at a fixed time or spatial interval.

```{r}
#| label: fig-post_dist
#| fig-cap: "Posterior distribution for models"
#| fig-subcap: ["Logistic Regression", "Poisson Regression", "Negative Binomial Regression"]
#| layout-ncol: 3
#| echo: false
#| warning: false
#| message: false

pp_check(logistic_model)

pp_check(poisson_model)

pp_check(neg_bio_model)
```


\newpage


# References

